---
title: "Journal 2"
author: "Bogdan Cuza"
---

```{r, globalsettings, echo=FALSE, warning=FALSE, results='hide'}
library(knitr)

knitr::opts_chunk$set(echo = TRUE)
opts_chunk$set(tidy.opts=list(width.cutoff=100),tidy=TRUE, warning = FALSE, message = FALSE,comment = "#>", cache=TRUE, class.source=c("test"), class.output=c("test2"))
options(width = 100)



colorize <- function(x, color) {sprintf("<span style='color: %s;'>%s</span>", color, x) }

```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(position = c('top', 'right'))
#klippy::klippy(color = 'darkred')
#klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```

Last compiled on `r format(Sys.time(), '%B, %Y')`

<br>

------------------------------------------------------------------------

# Lab 2 Networks as causes: Spatial regression analysis

------------------------------------------------------------------------

Loading packages:
```{r}
library(network)
library(igraph)
library(reshape2)
library(tidyverse)
```

Loading data:
```{r}
url1 <- "https://github.com/rensec/sasr08/raw/main/g_adj_matrix_simple.csv"
g_matrix <- read.csv(file = url1, header = FALSE)
g_matrix
```

Convert to matrix:
```{r}
g_matrix <- as.matrix(g_matrix)
g_matrix
```

```{r}
rownames(g_matrix) <- 1:nrow(g_matrix)
colnames(g_matrix) <- 1:ncol(g_matrix)
g_matrix
```

**ANSWER**: There are 5 nodes, and 4 edges.

Retrieve additional information
```{r}
g_nodes_age <- read.csv(file = "https://github.com/rensec/sasr08/raw/main/g_nodes_age.csv")
g_nodes_age
```

Convert to igraph object:
```{r}
g <- graph_from_adjacency_matrix(g_matrix)
class(g)
g
```

Add attributes and plot:
```{r}
g <- set_vertex_attr(g, 
                     name = "age",
                     value = g_nodes_age$age)
plot(g)
```

Also with edge list:
```{r}
g_elist <- read.csv(file = "https://github.com/rensec/sasr08/raw/main/g_elist.csv")
g_elist
g_from_elist <- graph_from_data_frame(g_elist) 
plot(g_from_elist)
```

**NOTE**: must include isolate

```{r}
g_from_elist <- graph_from_data_frame(g_elist, vertices = g_nodes_age) 
plot(g_from_elist)
```

From adjacency list to igraph object

```{r}
g_adj_list <- read.csv(file = "https://github.com/rensec/sasr08/raw/main/g_adj_list.csv")
g_adj_list
g_elist_from_alist <- g_adj_list %>% 
  select(id, friend1:friend2) %>% # keep only the network variables (and id)
  melt( id.vars = "id") # from the reshape2 package

g_elist_from_alist <- g_elist_from_alist%>% 
  filter(!is.na(value)) %>% # drop the missings
  rename(from ="id", to = "value", sourcevar= "variable") %>% #just nice for interpretation
  relocate(to, .after=from) #move around the columns
g_elist_from_alist
```

**ANSWER**: Keeping `sourcevar` is useful for maintaining the relatioship between the original data and the subsequent quantitative analysis. In other words, it might be important for the theory to know why a certain edge exists.

```{r}
nodelist <- select(g_adj_list,id,age)
g_from_alist <- graph_from_data_frame(g_elist_from_alist, vertices = nodelist) #This is a proper igraph graph object
plot(g_from_alist)
```

Conversions:

To adjacency matrix:

```{r}
g_adj_matrix <- as_adjacency_matrix(g, sparse = FALSE)
g_adj_matrix
```

To edge list:
```{r}
g_edgelist <- igraph::as_data_frame(g_from_elist, what = "edges") 
g_edgelist 
```

note: missing isolate

to get nodes:

```{r}
igraph::as_data_frame(g_from_elist, what = "vertices") 
```

# Knecht data

```{r}
library(igraph)
library(haven) # to read Stata files
library(reshape2)
library(tidyverse)
```

## load data:

```{r}
PupilsWaveV <- read_dta(file = "PupilsWaveV.dta")
class12b <- PupilsWaveV %>% 
  filter(schoolnr == "12b")  # keep only one class
```

## select variables and into igraph:
```{r}
class12b_clean <- class12b %>%
  select(namenr, age, sex, friend1:friend12)
  

class12b_elist <- class12b_clean %>%
  select(namenr, friend1:friend12) %>%
  melt(id.vars = "namenr") %>%
  filter(!is.na(value)) %>%
  rename(from = "namenr", to = "value", sourcevar = "variable") %>%
  relocate(to, .after = from)
```

## this is a directed network, but we can check using R:

```{r}
g_class12b <- graph_from_data_frame(class12b_elist, vertices = class12b_clean)
is_directed(g_class12b)
```

## plot:
```{r}
plot(g_class12b)
```

## five number summary:

### size:

```{r}
vcount(g_class12b)
ecount(g_class12b)
```

### density:
```{r}
edge_density(g_class12b)
```

### components:
```{r}
count_components(g_class12b)
```

### diameter:
```{r}
diameter(g_class12b)
mean_distance(g_class12b)
```

### clustering:
```{r}
transitivity(g_class12b, type="average")
```

## test the hypothesis that girls have higher degree than boys in this network
```{r}
degrees <- degree(g_class12b, mode = "all")

node_data <- data.frame(
  namenr = V(g_class12b)$name,
  sex = V(g_class12b)$sex,
  degree = degrees
)
```

```{r}
boxplot(degree ~ sex, 
        data = node_data, 
        main = "Degree Distribution by Gender",
        xlab = "Sex", 
        ylab = "Degree",
        col = c("lightblue", "pink"))
```

t test: 

```{r}
t_test_result <- t.test(degree ~ sex, data = node_data)
print(t_test_result)
```

We fail to reject the null hypothesis.

### final plot:

```{r}
node_colors <- ifelse(V(g_class12b)$sex == 1, "skyblue", "salmon")
node_sizes <- degree(g_class12b, mode = "all") * 2
plot(g_class12b, 
     vertex.color = node_colors, 
     vertex.size = node_sizes,
     vertex.label = NA,
     edge.arrow.size = 0.5,
     main = "Class 12b Friendship Network")
```