---
title: "Journal 2"
author: "Bogdan Cuza"
---

```{r, globalsettings, echo=FALSE, warning=FALSE, results='hide'}
library(knitr)

knitr::opts_chunk$set(echo = TRUE)
opts_chunk$set(tidy.opts=list(width.cutoff=100),tidy=TRUE, warning = FALSE, message = FALSE,comment = "#>", cache=TRUE, class.source=c("test"), class.output=c("test2"))
options(width = 100)



colorize <- function(x, color) {sprintf("<span style='color: %s;'>%s</span>", color, x) }

```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(position = c('top', 'right'))
#klippy::klippy(color = 'darkred')
#klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```

Last compiled on `r format(Sys.time(), '%B, %Y')`

<br>

------------------------------------------------------------------------

# Lab 2 Networks as causes: Spatial regression analysis

------------------------------------------------------------------------

Loading packages:
```{r}
library(network)
library(igraph)
library(reshape2)
library(tidyverse)
library(spdep)
```

## Loading data:
```{r}
data(columbus)
mydata <- columbus
attach(mydata)
neighbors <- col.gal.nb
listw <- nb2listw(neighbors, style = "W")
summary(listw)
```

## Linear Regression Analysis

```{r}
plot(neighbors, coords)
```

```{r}
olsreg <- lm(CRIME ~ INC + HOVAL)
summary(olsreg)
```

## Spatial Analysis
```{r}
moran.test(CRIME, listw)
```

```{r}
lm.LMtests(olsreg, listw, test=c("LMlag", "LMerr"))
```

**NOTE** Since we reject both hypotheses, we choose based either on theory or test them all and see what results they give.

### Spatial Log Model

```{r}
slm <- lagsarlm(CRIME ~ INC + HOVAL, data = mydata, listw, Durbin = "FALSE")
summary(slm)
```

question: Durbin model? 

I need to understand these paragraphs better:

> Note that we explicitly specify that this is not a Durbin model. Just like when using the linear regression model, we find a significantly negative effect of household income and housing values on the crime rate. Coefficients are very similar to the LRM coefficients. This is often the case given that LRM coefficients are technically still BLUE. The spatial parameter rho is positive and significant showing spatial dependence.
We now turn to the direct and indirect effects. For this purpose, we follow the approximation method by Lesage and Pase (2009) . We specify ‘zstats = TRUE’ to obtain p-values. We use a simulation with 200 iterations to obtain p-values.

```{r}
im<-impacts(slm, listw = listw, zstats=TRUE, R=200)
sums<-summary(im,  zstats=T)
data.frame(sums$res)

data.frame(sums$pzmat)
```

Note: how to interpret this? what does this do?